---
title: "BPHH2019 S04 Network Analysis"
output: html_notebook
---

### Intro

As a very short intro to network analyses of psychiatric symptoms, we are going to !!gratefully!! borrow some data from the network summer school in 2017 on PTSD symptom data. See this reference: 

Armour, C., Fried, E. I., Deserno, M.K., Tsai, J. & Pietrzak, R.H. (2017) A network analysis of DSM-5 posttraumatic stress disorder symptoms and correlates in U.S. military veterans. Journal of Anxiety Disorders, 45, 49-59. 
 	
### Set-up 	
 	
First you will need to install some packages:  

```{r, echo=TRUE, message=TRUE, warning=TRUE}
install.packages("qgraph")
install.packages("bootnet")
install.packages("dplyr")
```
and then add
```{r,}
library("qgraph")
library("dplyr")
library("bootnet")
```

### Get data

To get the data, go to this website: https://osf.io/8kgey
download the data, navigate to the directory
setwd("/Users/myname/data.....")

```{r}
setwd("/Users/henkcremers/Google Drive/Work/Teaching/uva_resmas_BPHH/2019:20/S04/CourseMaterials/Day 3/1 MRF 2/")
load("data.Rdata")
```


In order to make the output readable, you need to to add the symptoms names.

```{r, echo=TRUE, message=FALSE}
names<-c("B1","B2", "B3", "B4", "B5", "C1", "C2", "D1", "D2", "D3", "D4", "D5", "D6", "D7", "E1", "E2", "E3", "E4", "E5", "E6")
longnames <- c("Intrusive thoughts", "Nightmares", "Flashbacks", "Emotional cue reactivity", "Physiological cue reactivity", "Avoidance of thoughts", "Avoidance of reminders", "Trauma-related amnesia", "Negative beliefs", "Blame of self or others", "Negative trauma-related emotions", "Loss of interest", "Detachment", "Restricted affect", "Irritability/anger", "Self-destructive/reckless behavior", "Hypervigilance", "Exaggerated startle response", "Difficulty concentrating", "Sleep disturbance")
gr <- list("Intrusions"=c(1:5), "Avoidance"=c(6:7), "Cognition & mood alterations"=c(8:14),"Arousal & reactivity alterations"=c(15:20)) 
```

### and then some estimation and plotting - the fun stuff

And now estimate the network, and plot the data

```{r,}
# estimate network
n1 <- estimateNetwork(data, default="EBICglasso")


# plot network
g1<-plot(n1, labels=names, layout="spring", vsize=6, cut=0, border.width=1.5, border.color="black", 
         groups=gr, color=c('#a8e6cf', '#dcedc1', '#ffd3b6', '#ff8b94'), nodeNames = longnames,legend.cex=.4)
```

This is a visualization of the network structure (covariance) of the data. 
**What do you see?** Anything that seem noteworthy - just present. 


```{r}
c1 <- centralityPlot(n1,labels=longnames)
```
** Which symptoms have the highest centrality?** Does that make sense ? 

** What are some possible implications for highly central symptoms ? ** 

That is all (for now) - this was just meant to give you some idea on how to run a network analyses. Any of the methodological details are not important for now. Just try to think of the pro/cons of network analyses and how/if it could help the discussed issues in biological psychiatry. 

