---
title: "BPHH2019 S04 Genetics"
output: html_notebook
---

### Intro

You will work with the R package from PsyGeNET (Psychiatric disorders Gene association NETwork) go check this website: http://www.psygenet.org/web/PsyGeNET/menu/Rpackage.

This packages helps to get a sense of the evidence on the relation between a certain SNPs and a psychiatric disorders (gene-disease association; GDA). This is not the "definitive guide", but a helpful tool, with the some nice illustrations.

The instructions below are really just a simplified version of the manual, so for more detailed instructions and options, scroll down the page http://www.psygenet.org/web/PsyGeNET/menu/Rpackage, to the "vignette". I have listen some questions in blue, try to answer those. 

### Set-up 	
 	
First you will need to install the psygenet package:

for R version >3.6

```{r, echo=TRUE, message=TRUE, warning=TRUE}
install.packages("BiocManager")
BiocManager::install(c("psygenet2r"))
```

for R < 3.5

```{r}
source( "http://bioconductor.org/biocLite.R" )
biocLite( "psygenet2r" )
```

and then add

```{r,}
library("psygenet2r")
```

### load the genes you want to look at 

I have picked some of the (IMHO) most "hyped" genes in the psychiatry world, especially from the so called gene x environment interaction literature.

Serotonin transporter gene (SLC6A4)
Dopamine transporter gene DAT (SLC6A3)
catechol-O-methyltransferase (COMT)
brain derived neurotrophic factor (BDF)

But the are obviously many (>20.000) more (and each with many loci), and if you come across something in the literature, you can look the official code name here: https://www.genenames.org/ and add it to the list. But maybe for now, let start of with these genes:

```{r, echo=TRUE}
genesOfInterest <- c("SLC6A4","SLC6A3","COMT","BDNF")
```

Are the genes in the database? 

```{r}
m1 <- psygenetGene(
gene = genesOfInterest,
database = "ALL",
verbose = FALSE,
warnings = FALSE
)
```

### which disorders are associated with these genes? 

```{r}
plot(m1)
```

**Can you descirbe the pattern ? What does that say about the specificty of GDA's ?**


from these associations, you can see in which directions they go (increase/decrease)

```{r}
geneAttrPlot( m1, type = "evidence index" )
```

Next you can look at what the function is of those genes using the so called The PANTHER Protein Class Ontology. But this step delves a lot deeper into the biology that we are covering in this course. Skip if you like. 

```{r}
pantherGraphic( genesOfInterest, "ALL")
```

More relevant for this course, is the following: how strong is the evidence for these GDA's? The following plot option produces a heatmap (color coding the strength of evidence) of the GDA's. 


```{r}
plot(m1,type = "GDCA heatmap")
```
**What do you notice ?** If the values are fairly high, what would be a critical explanation? 
 
 
In case you a have a lot of time, you can go ahead and look at the papers that presented evidence for the association. Specifically, the following code pulls-up all the sentences from the papers that the discuss the GDA. The output is not that readable, but it might still be illustrative to look at the wording (how sure do the authors seem to be), and get a sense of the strength of the evidence, for instance for major depression.


```{r}
m2 <- psygenetGeneSentences(
geneList = genesOfInterest,
database = "ALL"
)
```


```{r}
sentences <- extractSentences( m2,
disorder = "major depressive disorder" )
sentences
```




